A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.

### 2020-02-05 paper.Rmd
implements suggestions from #109. Closes #109

@@

The README and data dictionary provides _human readable_ information on the data. To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata. This provides a structure allowing the data to be indexed and searched online, through services such as google datasets search [@google-data-search]. An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema]. For example, a dataset "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), and the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).

@@

Any code used to clean and tidy the raw data should be provided in the `data-raw` directory. Ideally this would involve only scripted languages, but if other practical steps were taken to clean up the data, these should be recorded in a plain text or markdown file.

@@

In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databases. This approach **gets the data somewhere rather than nowhere**. Its minimal features means it is simple to maintain. In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organize. A downside is that it does not scale to larger data. Self hosting the data is an option, but we discourage this, as it may succumb to bit rot - where data is corrupted, degraded, or servers turn off.

In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data. For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank), and so on.

@@

Tooling for producing data documentation information speeds up and simplifies the process of sharing data. Machine-readable metadata that can be indexed by google is created using the `dataspice` package [@dataspice]. To help create a data dictionary, we recommend the `codebook` package in R [@Arslan2019], which also generates machine readable metadata. Data dictionaries are implemented in other software such as STATA, which provides a "codebook" command. Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that allow  checksum file comparisons for versioning. Note that is different to Frictionless Data's tabular data package spec [@frictionlessdata-data-pkg].

@@

```{r sdss-spectra, fig.cap = "Spectra image of wavelength against flux from SDSS from two sources. Part A shows the example image taken from SDSS website. Part B shows the replicated figure using data provided as CSV, demonstrating great reproducibility.", out.width = "90%", fig.align = "center"}

>>>>

Correspondence to:

Nicholas Tierney (nicholas.tierney@gmail.com) & Karthik Ram (k@karthik.io)

@@

The README and data dictionary provides _human readable_ information on the data. To ensure data types are preserved - for example, dates are dates, names are characters - there needs to be some form of _machine readable_ metadata. This provides a structure allowing the data to be indexed and searched online, through services such as google datasets search [@google-data-search]. An excellent standard for metadata is Table Schema written by frictionless data [@Fowler2017]. For example, a dataset "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), and the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).

@@

Any code used to clean and tidy the raw data, should be provided in the `data-raw` directory. This cleaned up data, analysis ready data should be placed in the `data/` directory. Ideally this would involve only scripted languages, but if other practical steps were taken to clean up the data, these should be recorded in a plain text or markdown file.

@@

In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or in a serverless database. This approach **gets the data somewhere rather than nowhere**. Its minimal features means it is simple to maintain. In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organize. A downside is that it does not scale to larger data. Self hosting the data is an option, but we discourage this, as it may succumb to bit rot - where data is corrupted, degraded, or servers turn off.

In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data. For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank).

@@

Tooling for producing data documentation information speeds up and simplifies the process of sharing data. Machine-readable metadata that can be indexed by google is created using the `dataspice` package [@dataspice]. To help create a data dictionary, we recommend the `codebook` package in R [@Arslan2019], which also generates machine readable metadata. Data dictionaries are implemented in other software such as STATA, which provides a "codebook" command. Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that allow  checksum file comparisons for versioning. Note that is different to Frictionless Data's tabular data package spec [@Fowler2017].

@@

```{r sdss-spectra, fig.cap = "Visualisations from SDSS can be effecticely replicated. The spectra image of wavelength against flux from SDSS from two sources. Part A shows the example image taken from SDSS website. Part B shows the replicated figure using data provided as CSV - although there is some data cleaning or filtering since the numbers are not identical, it still demonstrates the great potential of reproducibility.", out.width = "90%", fig.align = "center"}
